An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2022-06-30 2022/OPTIMA/paper.tex

% interacttfssample.tex
% v1.05 - August 2017

\documentclass[]{interact}

\usepackage{color}
\usepackage{hyperref}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

%\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled
%\setlength\bibindent{2em}% To increase hanging indent in bibliography when line spacing is doubled

\usepackage[numbers,sort&compress]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{[}{]}{,}{n}{,}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

\begin{document}

%\articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria}

\author{
\name{Konstantin Barkalov, Vladimir Grishagin and Evgeny Kozinov\thanks{CONTACT Konstantin Barkalov. Email: konstantin.barkalov@itmm.unn.ru}}
\affil{Institute of Information Technology, Mathematics and Mechanics, Lobachevsky State University, 23 Gagarin ave., 603022 Nizhny Novgorod, Russia}
}

\maketitle

\begin{abstract}
The paper is devoted to consideration of multiextremal multicriteria black-box problems with time-consuming criteria and proposes a new computational algorithm for solving such problems. This algorithm is based on combination of convolution approach, dimensionality reduction, and information-statistical global optimization algorithm. For approximating the Pareto set a family of maximum convolutions is considered generating problems of scalar multiextremal optimization. The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by an accelerated information-statistical global optimization method. This method applies the technique of local refinements and uses in the course of current univariate optimization the information about criteria that was obtained when solving completed optimization problems. These features allow one to accelerate the search of global solution.  
The efficiency of the proposed algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems. Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.
\end{abstract}

\begin{keywords}
Multiobjective optimization; global optimization; dimensionality reduction; search information; nature-inspired metaheuristics; numerical comparison
\end{keywords}

\section{Introduction}

@@

If number of completed trials is not exactly divisible by $T$, where \textit{global-to-local ratio} $T$ is a predefined number, then
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q} + v_{q-1}}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

\begin{table}
\tbl{Results of the experiment with bi-criteria MCO problems of dimension 2.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \midrule
  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \bottomrule
\end{tabular}}
\label{tab1}
\end{table}

@@

\begin{table}
\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \midrule
  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \bottomrule
\end{tabular}}
\label{tab2}
\end{table}

@@

\begin{table}
\tbl{Results of testing 2-dimensional MCO problems with 4 criteria.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \midrule
  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \bottomrule
\end{tabular}}
\label{tab3}
\end{table}

@@

\section*{Disclosure statement}

No potential conflict of interest was reported by the authors.

\section*{Funding}

This work was supported by the Russian Science Foundation, project No. 21-11-00204.

%
% При загрузке статьи в журнальную систему - подставить библиографию из bbl-файла в tex-файл
%

\bibliographystyle{tfs}
\bibliography{bibliography}

\end{document}
+

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}

\usepackage{amsmath}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\begin{document}
%
\title{An accelerated algorithm for finding efficient solutions in multiobjective problems with black-box multiextremal criteria\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}}
%
\titlerunning{Algorithm for finding efficient solutions in multiobjective problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Vladimir Grishagin\orcidID{0000-0002-2884-3670} \and
Evgeny Kozinov\orcidID{0000-0001-6776-0096}}
%
\authorrunning{K. Barkalov, V. Grishagin and E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru}, \email{vagris@unn.ru}}

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The paper is devoted to consideration of multiextremal multicriteria black-box problems with time-consuming criteria and proposes a new computational algorithm for solving such problems. This algorithm is based on combination of convolution approach, dimensionality reduction, and information-statistical global optimization algorithm. For approximating the Pareto set a family of maximum convolutions is considered generating problems of scalar multiextremal optimization. The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by an accelerated information-statistical global optimization method. This method applies the technique of local refinements and uses in the course of current univariate optimization the information about criteria that was obtained when solving completed optimization problems. These features allow one to accelerate the search of global solution.  
The efficiency of the proposed algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems. Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.

\keywords{Multiobjective optimization \and Global optimization \and Dimensionality reduction \and Search information \and Nature-inspired metaheuristics \and Numerical comparison.}
\end{abstract}
%
%
%
\section{Introduction}

@@

If number of completed trials is not exactly divisible by $T$, where \textit{global-to-local ratio} $T$ is a predefined number, then
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{equation}\label{eq:18}
\begin{matrix}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q} + v_{q-1}}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{matrix}
\end{equation}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

\begin{table}
\caption{Results of the experiment with bi-criteria MCO problems of dimension 2.}
\centering
\begin{tabular}{ccccccc} \hline
 Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \hline
  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \hline
\end{tabular}
\label{tab1}
\end{table}

@@

\begin{table}
\caption{Results of testing bi-criteria 4-dimensional MCO problems.}
\centering
\begin{tabular}{ccccccc} \hline
 Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \hline
  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \hline
\end{tabular}
\label{tab2}
\end{table}

@@

\begin{table}
\caption{Results of testing 2-dimensional MCO problems with 4 criteria.}
\centering
\begin{tabular}{ccccccc} \hline
 Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \hline
  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \hline
\end{tabular}
\label{tab3}
\end{table}

@@

\bibliographystyle{splncs04}
\bibliography{bibliography}

\end{document}
